# Data Types
There are five basic data types in R:

- **Numeric** - This is the default treatment for numbers. This data type includes integers and doubles.
    - *Double* - A double allows you to store numbers as decimals. This is the default treatment for numbers.
    - *Integer* - An integer is a subset of the numeric data type. This type will only allow whole numbers and is denoted by the letter "L".
- **Complex** - This type is created by using the imaginary variable "i".
- **Character** - This type is used for storing non-numeric text data.
- **Logical** - Sometimes referred to as "boolean", this data type will store either "TRUE" or "FALSE".
- **Raw** - Used less often, this data type will store data as raw bytes.

## Numeric
### Double
```{r}
x <- 6.2
class(x)
typeof(x)
is.numeric(x)
is.integer(x)
is.double(x)
```
### Integer
```{r}
x <- 6L
class(x)
typeof(x)
is.numeric(x)
is.integer(x)
is.double(x)
```
## Complex
```{r}
x <- 6i
class(x)
typeof(x)
is.numeric(x)
is.integer(x)
is.double(x)
```
## Character
```{r}
x <- "Hello!"
class(x)
typeof(x)
```
## Logical
```{r}
x <- TRUE
class(x)
typeof(x)
```

## Raw
```{r}
x <- charToRaw("Hello!")
print(x)
class(x)
typeof(x)

x <- intToBits(6L)
typeof(x)
print(x)
class(x)
typeof(x)
```