# Data Structures

There are six basic data structures in R

- **Vectors** - Vectors contain of data which is ordered and of the same type.
- **Lists** - Lists are a collection of objects.
- **Matrices** - A matrix is a two-dimensional array where the data is all of the same type.
- **Factors** - Factors are used to designate levels within categorical data.
- **Data Frames** - A data frame contains two-dimensional data where the data can have different types.
- **Arrays** - Arrays are objects which have more than two dimensions (n-dimensional).

## Vectors
```{r}
x <- c(1, 3, 3, 7)

print(x)
```

## Lists
```{r}
first_name <- "John"
last_name <- "Smith"
favorite_numbers <- c(1, 3, 3, 7)

person <- list(first_name, last_name, favorite_numbers)

print(person)
```

## Matrices
```{r}
x <- matrix(
        c(1,3,3,7,1,3,3,7,1,3,3,7)
        , nrow = 3
        , ncol = 4
        , byrow = TRUE)

print(x)
```

## Factors
```{r}
x <- c("Red", "Blue", "Red", "Yellow", "Yellow")

colors <- factor(x)

print(colors)
```

## Data Frames
```{r}
people <- c("John", "Jane")
id <- c(1, 2)
df <- data.frame(id = id, person = people)

print(df)
```

## Arrays
```{r}
x <- array(
        c(1,3,3,7,1,3,3,7,1,3,3,7)
        , dim = c(1,4,3))

print(x)
```