# Exercises {.unnumbered}

## Questions

::: callout
## Exercise: 5-A

Write a function called "multiply" that will accept two numbers as arguments and will output the product of those two numbers when called as is demonstrated below.

``` r
multiply(3, 3)
# [1] 9
```
:::

::: callout
## Exercise: 5-B
Write an equation which will return the remainder of 12 divided by 8.
:::

::: callout
## Exercise: 5-C
Write an equation which will return the remainder of 36 divided by 10.
:::

::: callout
## Exercise: 5-D
Write a "while" loop which prints all even numbers from 0 to 10.

It's possible for this task to be accomplished in several ways; however, the output of your program should always look like this:

``` r
# [1] 0
# [1] 2
# [1] 4
# [1] 6
# [1] 8
# [1] 10
```
:::

::: callout
## Exercise: 5-E

You are given a vector that looks like this:

``` r
numbers <- c(0:12)
```

Write a for loop which loops through your vector and prints any element that is greater than or equal to 3.

It's possible for this task to be accomplished in several ways; however, the output of your program should always look like this:

``` r
# [1] 3
# [1] 4
# [1] 5
# [1] 6
# [1] 7
# [1] 8
# [1] 9
# [1] 10
# [1] 11
# [1] 12
```
:::

::: callout
## Exercise: 6-A

Convert the following character variable to a variable with the data type "raw":

``` r
x <- "Trevor rocks"
```

You should store your raw data in a variable named "raw_data", print the data to the console, and check the data type with the "typeof" function. Your output should look like the following:

``` r
print(raw_data)
# [1] 54 72 65 76 6f 72 20 72 6f 63 6b 73
typeof(raw_data)
# [1] "raw"
```

:::

::: callout
## Exercise: 6-B

Create a variable named "spending" and give it a value of 120. Then create a variable named "budget" and give it a value of 100. Next, check whether spending is greater than budget and store the resulting logical data in a variable named "over_budget". Finally, print the value of "over_budget" variable and check it's data type with the "typeof" function.

Your final output should look like this:

``` r
print(over_budget)
# [1] TRUE
typeof(over_budget)
# [1] "logical"
```

:::

::: callout
## Exercise: 7-A
Create a vector named "animal" and give it the following three values: "cow", "cat", "pig". Create a second vector named "sound" and give it the following three values: "moo", "meow", "oink". Finally, create a data frame named "animal_sounds" and assign each of these vectors to be a column.

After printing the resulting data frame to the console, you should get the following output:

``` r
#   animal sound
# 1    cow   moo
# 2    cat  meow
# 3    pig  oink
```
:::

## Answers

::: callout
## Answer: 5-A

One way you could accomplish this task is demonstrated in the following solution.

```{r}
multiply <- function(x, y) {
  return (x * y)
}

multiply(3, 3)
```
:::

::: callout
## Answer: 5-B

A remainder is referred to as "modulus" in programming. We can use the "%%" operator to accomplish this. For this example, the output of your equation should be 4.

```{r}
12 %% 8
```
:::

::: callout
## Answer: 5-C

A remainder is referred to as "modulus" in programming. We can use the "%%" operator to accomplish this. For this example, the output of your equation should be 6.

```{r}
36 %% 10
```
:::

::: callout
## Answer: 5-D
Here's one way you could write your while loop to achieve this output:

```{r}
i <- 0

while (i <= 10) {
  print(i)
  i <- i + 2
}
```
:::

::: callout
## Answer: 5-E
Here's one way you could write your for loop to achieve this output:

```{r}
numbers <- c(0:12)

for (number in numbers) {
  if (number >= 3) {
    print(number)
  }
}
```
:::

::: callout
## Answer: 6-A

You can accomplish this task through the "charToRaw" function.

```{r}
x <- "Trevor rocks"
raw_data <- charToRaw(x)
print(raw_data)
typeof(raw_data)
```
:::

::: callout
## Answer: 6-B

The following example demonstrates how you can accomplish this task.

```{r}
spending <- 120
budget <- 100
over_budget <- spending > budget
print(over_budget)
typeof(over_budget)
```
:::

::: callout
## Answer: 7-A

The following example demonstrates how you can accomplish this task.

```{r}
animal <- c("cow", "cat", "pig")
sound <- c("moo", "meow", "oink")
animal_sounds <- data.frame(animal = animal, sound = sound)
print(animal_sounds)
```
:::