# Working with APIs

API stands for Application Programming Interface. These sorts of tools are commonly used to programmatically pull data from a third party resource. This chapter demonstrates how one can begin to leverage these tools in their own workflows.

The following example uses the Helium API to return data about it's blockchain network.

## Install Packages

```{r}
#| eval: false
install.packages(c('httr', 'jsonlite'))
```

## Require Packages

```{r}
library('httr')
library('jsonlite')
```

## Make Request

Pass a URL into the 'GET' function and store the response in a variable called 'res'.

```{r}
#| eval: false
res = GET("https://api.helium.io/v1/stats")
print(res)
```

```{md}
Response [https://api.helium.io/v1/stats]
  Date: 2022-08-04 01:25
  Status: 200
  Content-Type: application/json; charset=utf-8
  Size: 922 B
```

## Parse & Explore Data

Use the 'fromJSON' function from the 'jsonlite' package to parse the response data and then print out the names in the resulting data set.

```{r}
#| eval: false
data = fromJSON(rawToChar(res$content))

names(data)
```

```{md}
[1] "data"
```

Go one level deeper into the data set and print out the names again.

```{r}
#| eval: false
data = data$data

names(data)
```

```{md}
[1] "token_supply"     "election_times"   "counts"           "challenge_counts" "block_times"
```

Alternatively, you can loop through the names as follows.

```{r}
#| eval: false
for (name in names(data)){print(name)}
```

```{md}
[1] "token_supply"
[1] "election_times"
[1] "counts"
[1] "challenge_counts"
[1] "block_times"
```

Get the 'token_supply' field from the data.

```{r}
#| eval: false
token_supply = data$token_supply

print(token_supply)
```

```{md}
[1] 124675821
```

## Adding Parameters to Requests

Add 'min_time' and 'max_time' as parameters on a different endpoint and print the resulting 'fee' data.

```{r}
#| eval: false
res = GET("https://api.helium.io/v1/dc_burns/sum",
    query = list(min_time = "2020-07-27T00:00:00Z"
                 , max_time = "2021-07-27T00:00:00Z"))

data = fromJSON(rawToChar(res$content))
fee = data$data$fee
print(fee)
```

```{md}
[1] 10112755000
```

## Adding Headers to Requests

Execute the same query as above except this time specify headers. This will likely be necessary when working with an API which requires an API Key.

```{r}
#| eval: false
res = GET("https://api.helium.io/v1/dc_burns/sum",
    query = list(min_time = "2020-07-27T00:00:00Z"
                 , max_time = "2021-07-27T00:00:00Z"),
    add_headers(`Accept`='application/json', `Connection`='keep-live'))

data = fromJSON(rawToChar(res$content))
fee = data$data$fee
print(fee)
```

```{md}
[1] 10112755000
```

## Resources

- Blog post by Trevor French: <https://medium.com/trevor-french/api-calls-in-r-136290ead81d>

### Helpful APIs

- Meta Graph API: <https://developers.facebook.com/docs/graph-api/>
- Twitter API: <https://developer.twitter.com/en/docs/twitter-api>
- NASA APIs: <https://api.nasa.gov/>
- Etherscan API: <https://etherscan.io/apis>
- Covalent API: <https://www.covalenthq.com/docs/api/#/0/0/USD/1>
- EDGAR APIs from the SEC: <https://www.sec.gov/edgar/sec-api-documentation>
- Weather API: <https://openweathermap.org/api>
- Helium API: <https://docs.helium.com/api/>